\speculake: Hiding Malware in Speculative Execution

### 2018-08-08 paper/architecture.tex
arch pass

\section{Architecture}
\speculake malware is comprised of two independent pieces: a payload program,
and a trigger. The payload, and some variants of the trigger, must be  installed
on the victim's computer(e.g. via trojan, remote exploit, or phishing). A running
payload performs innocuous measurements while waiting for the trigger to
become present.

If the trigger is a local program interacting with the payload
via the indirect branch predictor both programs must run on the same physical CPU. We 
note that this constraint is not a significant burden as most OS schedulers
will use round robin algorithms to assign tasks to processors---eventually 
the trigger and payload will execute co-resident on a single core.

@@

If the trigger is an input the program parses innocuous entries waiting 
for a series of crafted entries to train the direct branch predictor 
before a specially crafted trigger packet. The payload then forces the CPU
to speculate by stalling a bounds check. A local array is then speculatively 
overflowed redirecting program execution speculatively 
to an address within the payload program specified by the specially 
crafted trigger packet provided by the attacker.

@@

We further assume that the malware has a specific trigger that the analyst is
not privy to, and the adversary can influence. In our examples, this
trigger is often behavior exhibited by some other (potentially benign) process
running on the same system as the malware. As the adversary is able to control
when such a trigger is deployed (potentially remotely), the analyst will not be
able to observe or force this trigger to happen at will. We emphasize that while
this may also be true for existing trigger-based malware, analysts can often
reverse engineer the trigger out of the malware, for example by observing
control flow within the malware and using adaptive
fuzzers~\cite{afl,stephens2016driller} to generate inputs that explore other
execution paths of the binary. In contrast, \speculake malware's trigger depends
on behavior in seemingly unrelated binaries, influencing malware behavior
through channels that are effectively invisible to the analyst. As with typical
malware, the analyst may attempt to reverse engineer the trigger to reveal the
malware's behavior, but we will show (in Section~\ref{subsec:nested-spec}) how
to defeat this type of analysis.

@@

>>>>

\section{Architecture}
\speculake malware is comprised of two independent pieces: a payload program,
and a trigger. The payload, and some form of the trigger, must be installed
on the victim's computer (e.g. via trojan, remote exploit, or phishing). A running
payload performs innocuous operations while waiting for the trigger to
become present.

One form the trigger can take is another local program that interacts with the
payload via the indirect branch predictor. In this case, both programs must run
on the same physical CPU. We note that this constraint is not a significant
burden, as programs can either use \texttt{taskset}, or, if not available, run
multiple instances or wait for the OS scheduler to execute both programs on
the same core.

@@

The trigger can also take the form of a special input to the payload program,
rather than a separate program.
In this case, the payload program parses input data and performs innocuous
operations with it. Once the trigger input is provided, it causes the program to
\emph{speculatively} overflow a buffer, despite correct bounds checks in the program.
The speculative buffer overflow (described in
Section~\ref{subsec:spec-buffer-overflow}) causes the program to speculatively
execute at an address chosen by the trigger input and controlled by the
attacker.

@@

We further assume that the malware has a specific trigger that the analyst is
not privy to, and the adversary can influence. In our examples, this
trigger is often behavior exhibited by some other (potentially benign) process
running on the same system as the malware. As the adversary is able to control
when such a trigger is deployed (potentially remotely), the analyst will not be
able to observe or force this trigger to happen at will. We emphasize that while
this may also be true for existing trigger-based malware, analysts can often
reverse engineer the trigger out of the malware, for example by observing
control flow within the malware and using adaptive
fuzzers~\cite{afl,stephens2016driller} to generate inputs that explore other
execution paths of the binary. In contrast, \speculake malware's trigger
influences behavior of the payload program speculatively, making it effectively
invisible to the analyst. As with typical
malware, the analyst may attempt to reverse engineer the trigger to reveal the
malware's behavior, but we will show (in Section~\ref{subsec:nested-spec}) how
this type of analysis can be defeated.

@@

In this section, we will describe the trigger program variant, and discuss how
input data can be used a speculative trigger in
Section~\ref{sec:spec-buffer-overflow}.
